\begin{landscape}

\clearpage
\subsection{Dataset Description}
\label{app:dataset}

A comprehensive description of the dataset is provided in 
Table~\ref{tab:dataset_structure}, detailing each column's category, 
description, and objective.

\begin{longtable}{llp{8cm}p{8cm}p{1.5cm}}
    \caption{Stock Market Dataset}
    \label{tab:dataset_structure} \\
    \hline
    \textbf{Feature} & \textbf{Category} & \textbf{Description} & \textbf{Objective} & \textbf{Selected}  \\
    \hline\hline
    \endfirsthead

    \multicolumn{4}{c}{\textit{Continued from previous page}} \\
    \hline
    \textbf{Feature} & \textbf{Category} & \textbf{Description} & \textbf{Objective} & \textbf{Selected} \\
    \hline\hline
    \endhead

    \hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
    \endfoot

    \hline
    \endlastfoot
        \texttt{datetime} 
        & Date 
        & Timestamp for the stock data. 
        & Serves as the index for time-series forecasting.
        & Yes
        \\
        \texttt{open} 
        & Price  
        & The stock's opening price for the day. 
        & Helps analyze price gaps and trend formations. 
        & Yes
        \\
        \texttt{high} 
        & Price 
        & The highest price reached during the day. 
        & Indicates intraday volatility and resistance levels.
        & Yes
        \\
        \texttt{low} 
        & Price 
        & The lowest price recorded during the day. 
        & Shows intraday volatility and potential support levels. 
        & Yes
        \\
        \texttt{close} 
        & Price 
        & The stock's closing price for the day. 
        & Primary feature for forecasting future prices. 
        & Yes 
        \\
        \texttt{volume} 
        & Price 
        & Total number of shares traded in a day. 
        & Measures market activity and confirms price movements. 
        & Yes
        \\ \rowcolor{lightgray}
        \texttt{sma} 
        & Trend 
        & Simple Moving Average, smooths price fluctuations over a period. 
        & Identifies long-term trends and reduces price noise. 
        Was Removed because ema reacts faster to price changes compared to sma, which treats all past prices equally.
        & No 
        \\
        \texttt{ema} 
        & Trend 
        & Exponential Moving Average, giving more weight to recent prices. 
        & Detects trend direction and reversals faster than SMA. 
        & Yes
        \\ \rowcolor{lightgray}
        \texttt{wma} 
        & Trend 
        & Weighted Moving Average, emphasizing recent data more. 
        & Highlights recent price movements more effectively. When
          compared to wma, ema’s weighting decays exponentially, which is smoother and often less noisy.
        & No
        \\
        \texttt{macd} 
        & Trend 
        & Difference between two emas to detect trend strength. 
        & Identifies trend reversals and momentum shifts. 
        & Yes
        \\
        \texttt{macd\_hist} 
        & Momentum 
        & The MACD Histogram is one of the most powerful tools for visualizing momentum shifts in price trends. A positive histogram 
        represents bullish momentum while negative represents bearish.
        & Identifies momentum shifts. 
        & Yes
        \\ \rowcolor{lightgray}
        \texttt{macd\_signal} 
        & Trend 
        & The signal line is derived directly from the macd line using 
          an ema.
        & Removed because it's directly correlated with mac and 
          $macd\_hist=macd-macd\_signal$.
        & No 
        \\
        \texttt{rsi}
        & Momentum 
        & Measures the speed and magnitude of price movements (0-100 scale). 
        & Detects overbought (>70) and oversold (<30) conditions.
        & Yes
        \\
        \texttt{mom} 
        & Momentum 
        & Tracks the rate of price change over a set period. 
        & Measures trend acceleration and potential reversals. 
        & Yes
        \\
        \texttt{stoch\_slow\_k}
        & Momentum 
        & Current closing price relative to the recent low/high range 
        & Identifies overbought or oversold conditions; signals 
        potential reversals 
        & Yes 
        \\
        \texttt{stoch\_slow\_d}
        & Momentum 
        & 3-period SMA of the Stoch Slow \%K line 
        & Smooths the \%K line to confirm trend signals and reduce noise
        & Yes 
        \\
        \texttt{cci} 
        & Momentum 
        & Measures price deviation from the historical average. 
        & Detects overbought/oversold market conditions. 
        & Yes
        \\
        \texttt{adx}
        & Trend 
        & Measures the strength of a trend. 
        & Confirms whether a trend is strong enough to trade. 
        & Yes
        \\
        \texttt{roc}
        & Momentum 
        & Calculates the percentage price change over a given period. 
        & Measures trend strength and momentum shifts. 
        & Yes
        \\
        \texttt{willr}
        & Momentum
        & Measures the current closing price relative to the highest high
        over a lookback period. 
        & Identifies overbought or oversold levels; helps spot potential 
        trend reversals. 
        & Yes   
        \\
        \texttt{bbands\_middle\_band} 
        & Volatility 
        & The simple moving average (\texttt{sma}) of the closing price 
        & Represents the trend baseline; used as a smoothed reference 
          for price movement 
        & Yes 
        \\ \rowcolor{lightgray}
        \texttt{bbands\_upper\_band} 
        & Volatility 
        & Middle Band plus a multiple of the standard deviation 
        & Indicates potential overbought zone; used for breakout or
          mean-reversion strategies. Removed because its derived from 
          the \texttt{bbands\_middle\_band} with a standard deviation.
        & No 
        \\ \rowcolor{lightgray}
        \texttt{bbands\_lower\_band}  
        & Volatility 
        & Middle Band minus a multiple of the standard deviation 
        & Indicates potential oversold zone; used for breakout or 
          mean-reversion strategies. Removed because its derived from 
          the \texttt{bbands\_middle\_band} with a standard deviation.
        & No 
        \\
        \texttt{atr} 
        & Volatility 
        & Measures market volatility based on price range expansions. 
        & Helps set stop-loss levels and detect risk conditions. 
        & Yes
        \\
        \texttt{obv}
        & Volume 
        & Cumulative volume flow based on price movements. 
        & Confirms trends using volume-based signals. 
        & Yes
        \\
        \texttt{mfi}
        & Volume
        & Combines price and volume to assess money flow. 
        & Detects buying and selling pressure for trend confirmation. 
        & Yes\\
        \hline
\end{longtable}
\end{landscape}
